A performance assessment of evolutionary algorithms in volunteer computing environments: the importance of entropy

### 2017-01-30 EvoPar2017/paper.Rnw
Prepares final version

% \author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
%   Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}

% \institute{Dept. of Computer Architecture and Technology, University
% of Granada, Spain \and
% Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}

\author{A.U. Thor \and A. N. Otherauthor \and Just A. Nother \inst{1}
  \and Jet A. Nother \inst{2}}

\institute{Invisible University, Frantzia \and
Institute of Higher Studies, Zootropa}

@@

Using this approach for creating distributed experiments is called {\em
  volunteer}, {\em cycle-scavenging}, or {\em opportunistic} computing
\cite{sarmenta2001volunteer} and it dates back, in different shapes
and underlying mechanisms, to the origin of the web \cite{david-seti:home}. Our interest
here, however, is to use it as a resource for evolutionary
computation, as our group has done for a long time \cite{jj-ppsn98anonymous}.

In this line of research that uses volunteer computing for
evolutionary algorithms, there are several pending issues. The first
and maybe most important is approaching volunteer computing as a
socio-technical system \cite{vespignani2009predicting} which integrates
user decisions and behavioral patterns in the system model; this
includes trying to optimize the number of users in a particular
experiment. The second line of research, although related to the
first, is more focused on the evolutionary algorithm itself and how
different design decisions will affect its performance.
% Is it ok that we say that the first line of research is the most important... but then we focus on the second?
We have approached the first issue in our previous work \cite{jj-ppsn98anonymous},
% I have justified to focus, in this case, on the second
but in this paper our focus will be in the second aspect: we will try to
design a decentralized system that, at the same time, is able to use
all available resources for finding the solution of an evolutionary algorithm. This design will be done incrementally by
changing client and the server and measuring its
impact on the overall performance: time and evaluations needed to find
the solution. Eventually, we want to find a system that, whatever the
number of users available to perform the experiment, is able to
maximize their contribution to the evolutionary algorithm, at the same
time that the evolutionary algorithm itself makes the most of those
contributions and is able to find
the solution to the problem in a minimum time, with the least
number of contributions.

@@

Volunteer computing involves users deciding to run a program that acts
as a client or as a peer in an experiment
and, as such, has been deployed in many different ways from the
beginning of the Internet, starting with the SETI@home framework for
processing extraterrestrial signals \cite{david-seti:home}, or a 
high-throughput queuing system such as HTCondor \cite{HTCondor}.
However,
the dual facts of the introduction of JavaScript as a universal language for the
browser and the browser itself as both an ubiquitous web and Internet client has
made this combination the most popular for volunteer computing
frameworks such as the one we are using here, and whose first version
was described in \cite{DBLP:conf/gecco/GuervosG15anonymous}. Systems based on the
JavaScript/browser combination
emphasize the ephemerality, ease of use, and universality, while
systems such as HTCondor or BOINC might be more adequate for work that % citation for BOINC (?)
require higher availability of volunteer resources, achieving this
with downloadable clients which in some cases, like HTCondor, need
administration access to the resources.

@@

In this section we present {\sf NodIO}, a distributed volunteer-based 
evolutionary computation system implemented in JavaScript. 
All parts of the framework are free and available
with a free license from \url{https://ANONYMOUS-URL}.
%\url{https://github.com/JJ/splash-volunteer}.

@@

Figure \ref{fig:system} describes the general system architecture; the 
algorithm is described in \ref{algorithm}. Different JavaScript libraries, such as JQuery or {\tt
  Chart.js} have been used to build the user interface elements of the
framework, which should be running in \url{http://ANONYMOUS-URL}, 
%\url{http://nodio-jmerelo.rhcloud.com}, 
a free resource hosted in the
OpenShift Platform as a Service\footnote{It is not guaranteed to be running,
or running the same version, when you read this, however; you can always get
the sources from GitHub and set it up yourself.}.
%
\begin{figure}[!t]
\centering
\includegraphics[width=3in]{system.pdf}
\caption{Description of the NodIO system. Clients execute a JavaScript EA
  in the browser, which, every 100 generations, sends the best
  individual and receives a random one back from the server.}
\label{fig:system}
\end{figure}

@@

This is one of the changes we intend to explore in the future, which
holds many challenges, such as gathering more users and improving the
efficiency of the algorithm run in an asynchronous way by the
users. In the first case, it will be interesting to make it more {\em
  social} by comparing local performance to other's contributions, and
showing it in a graph or being able to tweet advances or the fact that
the solution has been found. Anything that keeps the users running the
system and attracts new users will contribute to the speed of finding
the solution. And the algorithm will have to be improved, mainly by
making the population more diverse. Right now there is a rigid policy
of contribution to the pool: send the best individual, but this could
be relaxed. Some other policies could also be tested, as well as
running heterogeneous algorithms in the same way that was done in
\cite{DBLP:journals/grid/ValdezTGVO15anonymous}. Diversity, as in any other
evolutionary algorithm, seems to be the key, but there are many ways of
approaching it.

All these avenues of experimentation will be done openly following the
Open Science policy of our group, which, in fact, contributes to
establish trust and security between us and volunteers and is an
essential feature of the system. That is why this paper, as well as
the data and processing scripts, are published with a free license in GitHub at
\url{ANONYMOUS-URL}.
%\url{https://git.io/gecco-es-15}.

@@

% This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
% Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
% 2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
% gathering users for the new experiments.

Several projects\\
Taking\\
This\\
Much\\
Space\\

>>>>

\author{Juan-J.~Merelo\inst{1} \and Paloma de las Cuevas\inst{1} \and Pablo
  Garc\'ia-S\'anchez \inst{2} \and Mario Garc\'ia-Valdez \inst{3}}

\institute{Dept. of Computer Architecture and Technology, University
of Granada, Spain and CITIC, \email{(jmerelo,palomacd)@ugr.es} 
\and
Department of Computer Engineering, University of C\'adiz (Spain), \email{pablo.garciasanchez@uca.es}
\and
Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana, \email{amherag,mario@tectijuana.edu.mx}}

%% \author{A.U. Thor \and A. N. Otherauthor \and Just A. Nother \inst{1}
%%   \and Jet A. Nother \inst{2}}

%% \institute{Invisible University, Frantzia \and
%% Institute of Higher Studies, Zootropa}

@@

Using this approach for creating distributed experiments is called {\em
  volunteer}, {\em cycle-scavenging}, or {\em opportunistic} computing
\cite{sarmenta2001volunteer} and it dates back, in different shapes
and underlying mechanisms, to the origin of the web \cite{david-seti:home}. Our interest
here, however, is to use it as a resource for evolutionary
computation, as our group has done for a long time \cite{jj-ppsn98}.

In this line of research that uses volunteer computing for
evolutionary algorithms, there are several pending issues. The first
and maybe most important is approaching volunteer computing as a
socio-technical system \cite{vespignani2009predicting} which integrates
user decisions and behavioral patterns in the system model; this
includes trying to optimize the number of users in a particular
experiment. The second line of research, although related to the
first, is more focused on the evolutionary algorithm itself and how
different design decisions will affect its performance.
% Is it ok that we say that the first line of research is the most important... but then we focus on the second?
We have approached the first issue in our previous work \cite{jj-ppsn98},
% I have justified to focus, in this case, on the second
but in this paper our focus will be in the second aspect: we will try to
design a decentralized system that, at the same time, is able to use
all available resources for finding the solution of an evolutionary algorithm. This design will be done incrementally by
changing client and the server and measuring its
impact on the overall performance: time and evaluations needed to find
the solution. Eventually, we want to find a system that, whatever the
number of users available to perform the experiment, is able to
maximize their contribution to the evolutionary algorithm, at the same
time that the evolutionary algorithm itself makes the most of those
contributions and is able to find
the solution to the problem in a minimum time, with the least
number of contributions.

@@

Volunteer computing involves users deciding to run a program that acts
as a client or as a peer in an experiment
and, as such, has been deployed in many different ways from the
beginning of the Internet, starting with the SETI@home framework for
processing extraterrestrial signals \cite{david-seti:home}, or a 
high-throughput queuing system such as HTCondor \cite{HTCondor}.
However,
the dual facts of the introduction of JavaScript as a universal language for the
browser and the browser itself as both an ubiquitous web and Internet client has
made this combination the most popular for volunteer computing
frameworks such as the one we are using here, and whose first version
was described in \cite{DBLP:conf/gecco/GuervosG15}. Systems based on the
JavaScript/browser combination
emphasize the ephemerality, ease of use, and universality, while
systems such as HTCondor or BOINC might be more adequate for work that % citation for BOINC (?)
require higher availability of volunteer resources, achieving this
with downloadable clients which in some cases, like HTCondor, need
administration access to the resources.

@@

In this section we present {\sf NodIO}, a distributed volunteer-based 
evolutionary computation system implemented in JavaScript. 
All parts of the framework are free and available
with a free license from \url{https://github.com/JJ/splash-volunteer}.

@@

Figure \ref{fig:system} describes the general system architecture; the 
algorithm is described in \ref{algorithm}. Different JavaScript libraries, such as JQuery or {\tt
  Chart.js} have been used to build the user interface elements of the
framework, which should be running in \url{http://nodio-jmerelo.rhcloud.com}, 
a free resource hosted in the
OpenShift Platform as a Service\footnote{It is not guaranteed to be running,
or running the same version, when you read this, however; you can always get
the sources from GitHub and set it up yourself.}.
%
\begin{figure}[!t]
\centering
\includegraphics[width=3in]{system.pdf}
\caption{Description of the NodIO system. Clients execute a JavaScript EA
  in the browser, which, every 100 generations, sends the best
  individual and receives a random one back from the server.}
\label{fig:system}
\end{figure}

@@

This is one of the changes we intend to explore in the future, which
holds many challenges, such as gathering more users and improving the
efficiency of the algorithm run in an asynchronous way by the
users. In the first case, it will be interesting to make it more {\em
  social} by comparing local performance to other's contributions, and
showing it in a graph or being able to tweet advances or the fact that
the solution has been found. Anything that keeps the users running the
system and attracts new users will contribute to the speed of finding
the solution. And the algorithm will have to be improved, mainly by
making the population more diverse. Right now there is a rigid policy
of contribution to the pool: send the best individual, but this could
be relaxed. Some other policies could also be tested, as well as
running heterogeneous algorithms in the same way that was done in
\cite{DBLP:journals/grid/ValdezTGVO15}. Diversity, as in any other
evolutionary algorithm, seems to be the key, but there are many ways of
approaching it.

All these avenues of experimentation will be done openly following the
Open Science policy of our group, which, in fact, contributes to
establish trust and security between us and volunteers and is an
essential feature of the system. That is why this paper, as well as
the data and processing scripts, are published with a free license in GitHub at \url{https://git.io/gecco-es-15}.

@@

This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
Ministry of Economy and Competitivity). We are also grateful to {\tt @otisdriftwood} for his help
gathering users for the new experiments.
